\propertitle

### 2019-11-27 maxNt_v6.tex
update

\pdfoutput=1
%% Author: PGL  Porta Mana
%% Created: 2015-05-01T20:53:34+0200
%% Last-Updated: 2019-11-26T22:02:42+0100
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newif\ifarxiv
\arxivfalse
\ifarxiv\pdfmapfile{+classico.map}\fi
\newif\ifafour
\afourfalse% true = A4, false = A5
\newif\iftypodisclaim % typographical disclaim on the side
\typodisclaimtrue
\newcommand*{\memfontfamily}{zplx}
\newcommand*{\memfontpack}{newpxtext}
\documentclass[\ifafour a4paper,12pt,\else a5paper,10pt,\fi%extrafontsizes,%
onecolumn,oneside,article,%french,italian,german,swedish,latin,
british%
]{memoir}
\newcommand*{\updated}{\today}
\newcommand*{\firstdraft}{4 November 2015}
\newcommand*{\firstpublished}{***}
\newcommand*{\propertitle}{Maximum-entropy distributions for a 
  neuronal population\\ from subpopulation data {\large\color{myred}[draft]}%\\{\large ***}%
}
\newcommand*{\pdftitle}{\propertitle}
\newcommand*{\headtitle}{Maximum-entropy from sample data}
\newcommand*{\pdfauthor}{P.G.L.  Porta Mana, Y. Roudi, V. Rostami, E. Torre}
\newcommand*{\headauthor}{Porta Mana \etal}
\newcommand*{\reporthead}{}% Report number

@@

The crucial point is our belief distribution for the activity of the
sample conditional on the activity of the larger population, given by
formula~\eqref{eq:conditional_hypergeometric} in
appendix~\ref{sec:derivation_maxent_fullnet}, reproduced here:
\begin{equation}
  \labelbis{eq:conditional_hypergeometric}
  \yG_{\ya\yA} \defd \pf(\ya \|\yA, n,N)=
  \binom{n}{\ya}\binom{N-n}{\yA-\ya}\binom{N}{\yA}^{-1}
\end{equation}
which leads to the equality of factorial
moments~\eqref{eq:equal_m_factorialmoments}, on which our approach rests.
The conditional probability above is characteristic of \enquote{drawing
  without replacement}
\citep[\chap~3]{jaynes1994_r2003}[\sect~4.8.3]{ross1976_r2010}[\sect~II.6]{feller1950_r1968}.
In other words, consider the pool of neurons which we equally believe could
have been recorded. $N$ is the size of that pool. Whether the equality of
such beliefs is justified, or for which sampling procedures it could be
justified, is a fundamental, deep question of sampling theory, which we
cannot discuss here of course \citep[see \eg\ the discussions, reviews, and
references in][]{ericson1969,smith1976}. $N$ therefore depends on factors
such as: the shape, dimensions, and technical specifications of the
recording probe; the inaccuracy in the insertion of the probe, leading for
example to slightly different insertion points or angles; the density of
neurons around the probe; and similar. $N$ can only be assessed case by
case. The plots of \fig~\ref{fig:plots_5mom} suggest that its order of
magnitude might be enough for qualitative inferences. As discussed in the
next section, even if the exact number $N$ were known, the maximum-entropy
distribution ought to be interpreted qualitatively or semi-quantitatively.

>>>>

\pdfoutput=1
%% Author: PGL  Porta Mana
%% Created: 2015-05-01T20:53:34+0200
%% Last-Updated: 2019-11-27T11:44:52+0100
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newif\ifarxiv
\arxivfalse
\ifarxiv\pdfmapfile{+classico.map}\fi
\newif\ifafour
\afourfalse% true = A4, false = A5
\newif\iftypodisclaim % typographical disclaim on the side
\typodisclaimtrue
\newcommand*{\memfontfamily}{zplx}
\newcommand*{\memfontpack}{newpxtext}
\documentclass[\ifafour a4paper,12pt,\else a5paper,10pt,\fi%extrafontsizes,%
onecolumn,oneside,article,%french,italian,german,swedish,latin,
british%
]{memoir}
\newcommand*{\updated}{\today}
\newcommand*{\firstdraft}{4 November 2015}
\newcommand*{\firstpublished}{***}
\newcommand*{\propertitle}{Maximum-entropy distributions for a 
  neuronal population\\ from subpopulation data {\large\color{myred}[draft]}%\\{\large ***}%
}
\newcommand*{\pdftitle}{\propertitle}
\newcommand*{\headtitle}{Maximum-entropy from sample data}
\newcommand*{\pdfauthor}{P.G.L.  Porta Mana, Y. Roudi, V. Rostami, E. Torre}
\newcommand*{\headauthor}{Porta Mana \etal}
\newcommand*{\reporthead}{}% Report number

@@

The crucial point is our belief distribution for the activity of the
sample conditional on the activity of the larger population, given by
formula~\eqref{eq:conditional_hypergeometric} in
appendix~\ref{sec:derivation_maxent_fullnet}, reproduced here:
\begin{equation}
  \labelbis{eq:conditional_hypergeometric}
  \yG_{\ya\yA} \defd \pf(\ya \|\yA, n,N)=
  \binom{n}{\ya}\binom{N-n}{\yA-\ya}\binom{N}{\yA}^{-1}
\end{equation}
which leads to the equality of factorial
moments~\eqref{eq:equal_m_factorialmoments}, on which our approach rests.
The conditional probability above is characteristic of \enquote{drawing
  without replacement}
\citep[\chap~3]{jaynes1994_r2003}[\sect~4.8.3]{ross1976_r2010}[\sect~II.6]{feller1950_r1968}.
In other words, consider the pool of all neurons which we believe -- with
equal degree for each neuron -- could have been recorded. $N$ is the size
of that pool. $N$ therefore depends on factors such as: the shape,
dimensions, and technical specifications of the recording probe; the
inaccuracy in the insertion of the probe, leading for example to slightly
different insertion points or angles; the density of neurons around the
probe. But it also depends on the homogeneity of the brain region where the
recording was made: if we believe that it doesn't matter whether the probe
had been inserted in some other point of the same brain region, then
formula~\eqref{eq:conditional_hypergeometric} is appropriate. Thus $N$ can
only be assessed case by case.

Whether such beliefs are justified, or for which sampling procedures they
can be justified, is a fundamental, deep question of sampling theory, which
we cannot discuss here \citep[see \eg\ the discussions, reviews, and
references in][]{ericson1969,smith1976,kruskaletal1979c,kruskaletal1980}.
It is of course possible to probabilistically assess, case by case, whether
this assumption is appropriate. It good to keep in mind that such
assessment rests, in turn, on similar assumptions at a higher level.

Luckily the plots of \fig~\ref{fig:plots_5mom} suggest that the order of
magnitude of $N$ ought to be enough for qualitative inferences. As
discussed in the next section, even if the exact number $N$ were known, the
maximum-entropy distribution ought to be interpreted qualitatively or
semi-quantitatively.
